The state learner -- a super learner for right-censored data

### 2024-05-14 tex/jrss-version/state-learner-jrss-style.tex
better naming of figure files

  & \qquad

@@

\begin{corollary}
  \label{cor:oracle-prop}
  For all \(P_0\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    \\
    & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

@@

An example of a targeted estimator is the one-step estimator, defined as
\begin{equation}
    \label{eq:one-step-def}
    \hat{\Psi}_{\text{OS}}(\data_n)
    =
    \tilde{\Psi}^0(\hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
    H(\data_n))
    + \empmeas{[\psi(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
      \hat{b}_n(\data_n) )]},
\end{equation}
where \( \empmeas \) is the empirical measure of a data set
\(\{O_i\}_{i=1}^n\).
Under suitable regularity conditions
we have the
following asymptotic expansion of the one-step estimator
\citep{pfanzagl1985contributions,van2003unified,fisher2021visually,kennedy2022semiparametric,rytgaard2022continuous},
\begin{equation*}
  \hat{\Psi}_{\text{OS}}(\data_n)- \Psi(P)
  =  \empmeas{[\psi(\blank ; \Lambda_1, \Lambda_2, \Gamma)]}
  +\mathrm{Rem}{(\hat{\Lambda}_{1,n},\hat{\Lambda}_{2,n},  \hat{\Gamma}_n, P)} + \smallO_{P}(n^{-1/2}),
\end{equation*}
where the remainder term has the form
\begin{equation}
  \label{eq:4}
  \mathrm{Rem}{(\hat{\Lambda}_{1,n},\hat{\Lambda}_{2,n},  \hat{\Gamma}_n, P)}
  = \mathcal{O}_P{
    \left\{
      \|\Lambda_1-\hat{\Lambda}_{1,n}\|^2
      +
      \|\Lambda_2-\hat{\Lambda}_{2,n}\|^2
      +
      \|\Gamma-\hat{\Gamma}_{n}\|^2
    \right\}
  },
\end{equation}
for some suitable norm \( \|\blank \| \), for instance the
\( \mathcal{L}_{P}^2 \)-norm. When equation~(\ref{eq:4}) holds and the
nuisance parameters $\Lambda_1$, $\Lambda_2$, and $\Gamma$ are
consistently estimated at rate \( \smallO_P{(n^{-1/4})} \), then
\begin{equation}
  \label{eq:3}
  \sqrt{n}(\hat{\Psi}_{\text{OS}}(\data_n)- \Psi(P)) \weakly \mathcal{N}(0,
  P{[\psi(\blank; \Lambda_1, \Lambda_2, \Gamma)^2]}),
\end{equation}
where we use \( \weakly \) to denote weak convergence \citep{van2000asymptotic}.
In particular, equation~(\ref{eq:3}) and Slutsky's lemma imply that we can
obtain asymptotically valid \((1-\alpha)\cdot100\%\) confidence intervals by
calculating
\begin{equation*}
  \left[
    \hat{\Psi}_{\text{OS}}(\data_n) - q_{\alpha/2} \hat{\sigma}(\data_n) ,
    \;
    \hat{\Psi}_{\text{OS}}(\data_n) + q_{\alpha/2} \hat{\sigma}(\data_n)
  \right],
\end{equation*}
where \( q_{\alpha} \) is the \( (1-\alpha) \)-quantile of the standard normal
distribution and
\( \hat{\sigma}(\data_n)^2 = \empmeas{ \left[ \psi(\blank;
    \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n), \hat{b}_n(\data_n))^2
  \right]} \).

@@

We compare the state learner to two IPCW based super learners: The
first superlearner, called IPCW(Cox), uses a Cox model with main
effects to estimate the censoring probabilities, while the second
super learner, called IPCW(KM), uses the Kaplan-Meier estimator to
estimate the censoring probabilities. The Cox model for the censoring
distribution is correctly specified in both simulation settings while
the Kaplan Meier estimator only estimates the censoring model
correctly in the simulation setting where censoring is
independent. Both IPCW super learners are fitted using the
\texttt{R}-package \texttt{riskRegression}
\citep{Gerds_Ohlendorff_Ozenne_2023}.
%
% | time | sim_setting | true_events | true_cens | at_risk |
% |------+-------------+-------------+-----------+---------|
% |   36 | original    |      24.619 |    61.853 |  25.774 |
% |   36 | indep_cens  |      24.674 |    38.740 |  46.141 |
%
The IPCW super learners use the integrated Brier score up to a fixed time
horizon (36 months). The marginal risk of the event before this time horizon is
\(\sim 24.6\)\%. Under the ``dependent censoring'' setting the marginal censoring
probability before the time horizon is \(\sim 61.9\)\%. Under the ``independent
censoring'' setting the marginal censoring probability before this time horizon
is \(\sim 38.7\)\%.

@@

In this section we use the prostate cancer data of
\cite{kattan2000pretreatment} to illustrate the use of the state
learner in the presence of competing risks. We have introduced the
data in Section~\ref{sec:numer-exper}. The data consists of 1,042
patients who are followed from start of followup until tumor
recurrence, death without tumor recurrence or end of followup
(censored) whatever came first. For the sole purpose of illustration,
we estimate the average treatment effect of hormone therapy on death
and tumor recurrence. To do this we adapt the estimation strategy of
Section~\ref{sec:targeted-learning} as follows.  We use the state
learner to rank libraries of learners for the cause-specific
cumulative hazard functions of tumor recurrence, death without tumor
recurrence, and censoring.  The libraries of learners include the five
learners described in Table~\ref{tab:zel-library}.  We use the same
set of learners to learn the cumulative hazard function of tumor
recurrence \( \Lambda_1 \), the cumulative hazard function of death
without tumor recurrence \( \Lambda_2 \), and the cumulative hazard
function of the conditional censoring distribution $\Gamma$. 
We then use the highest ranked combination
of learners and apply formula \ref{eq:one-step-comp-ate}.

@@

\begin{figure}
  \centering %
  \includeFigCond[width=.9\linewidth]{zelefski-real-data.pdf}
  \caption[]{The results of applying the 125 combinations of learners to the
    prostate cancer data set. The learners are \texttt{KM} (KM), \texttt{Cox
      strata CS} (strata), \texttt{Lasso} (lasso), \texttt{Elastic} (elastic),
    and \texttt{RF} (RF) as described Table~\ref{tab:zel-library}. The error
    bars are based on five repetitions using different splits. We refer to
    learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as `Tumor
    learner', `Death learner', and `Censoring learner', respectively.}
  \label{fig:zelefski-real}
\end{figure}

\begin{table}
  \caption{\label{tab:zelefski-real}The 10 best performing models in terms of integrated Brier score. The
    reported standard errors are based on five repetitions using different
    splits. The models are described in Table~\ref{tab:zel-library}. We refer to
    learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as `Tumor
    learner', `Death learner', and `Censoring learner', respectively.}
  \centering
  \input{./zel-tab.tex}
\end{table}

@@

\begin{figure}
  \centering%
  \includeFigCond[width=.8\linewidth]{zelefsky-data-target-par.pdf}
  \caption[]{Estimates of the average treatment effect of hormone therapy on the
    risk of tumor recurrence and death obtained from the prostate cancer study
    analyzed by \cite{kattan2000pretreatment}. The estimates are based on the
    estimator defined in equation~(\ref{eq:one-step-comp-ate}) with point wise
    95\% confidence intervals calculated as described in
    Section~\ref{sec:targeted-learning}. The estimates of the nuisance
    parameters are provided by the state learner.}
  \label{fig:zelefski-real-target}
\end{figure}

@@

\begin{proof}[Proof of Proposition~\ref{prop:stric-prop}]
  The result follows from Lemma~\ref{lemma:norm}.
\end{proof}

In the following, let $\Theta$ denote the function space consisting of all
conditional state occupation probability functions for some measure \( P \) as
defined in equation~(\ref{eq:F-def}).

\begin{proof}[Proof of Corollary~\ref{cor:oracle-prop}]
  First note that minimising the loss \( \bar{B}_{\tau} \) is equivalent to
  minimising the loss \( \bar{B}_{\tau,0} \), so the discrete super learner and
  oracle according to \( \bar{B}_{\tau} \) and \( \bar{B}_{\tau,0} \) are
  identical. By Lemma~\ref{lemma:norm}, \( R_0(F) \geq 0 \) for any
  \( F \in \Theta \), and so using Theorem 2.3 from \citep{vaart2006oracle} with
  \( p=1 \), we have that for all \( \delta >0 \),
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad + (1+\delta) \frac{16 K}{n}
    \log(1 + |\mathcal{F}_n|)\sup_{F \in \Theta}
    \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
    \frac{1}{\delta} + 1
    \right)
    \right\}
\end{align*}
where for each \( F \in \Theta \), \( (M(F), v(F)) \) is some Bernstein pair for
the function \(o \mapsto \bar{B}_{\tau,0}(F, o) \). As
\( \bar{B}_{\tau,0}(F, \blank) \) is uniformly bounded by \( \tau \) for any
\( F \in \Theta \), it follows from section 8.1 in \citep{vaart2006oracle} that
\( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) is a Bernstein pair for
\( \bar{B}_{\tau,0}(F, \blank) \). Now, for any \( a,b,c \in \R \) we have
\begin{align*}
  (a-c)^2 - (b-c)^2
  & = (a-b+b-c)^2 - (b-c)^2
  \\
  & = (a-b)^2 + (b-c)^2 +2(b-c)(a-b) - (b-c)^2
  \\
  & = (a-b)
    \left\{
    (a-b) +  2(b-c)
    \right\}
  \\
  & = (a-b)
    \left\{
     a + b -2c
    \right\},
\end{align*}
so using this with \( a=F(t, j, x) \), \( b=F_0(t, j, x) \), and
\( c = \1{\{\eta(t) = j\}} \), we have by Jensen's inequality
\begin{align*}
  & P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}
  \\
  & \leq
    2\tau\E_{P_0}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left\{
    \left(
    F(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    -
    \left(
    F_0(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    \right\}^2
    \diff t 
    \right]}
  \\
  & =2\tau
    \E_{P_0}\Bigg[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_0(t, j, X)
    \right)^2
  \\
  & \quad \quad \quad\quad \quad \quad \times
    \left\{
    F(t, j, X) +  F_0(t, j, X)-2 \1{\{\eta(t) = j\}}
    \right\}^2
    \diff t 
    \Bigg]
  \\
  & \leq
    8\tau \E_{P_0}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_0(t, j, X)
    \right)^2
    \diff t 
    \right]}.
  \\
  & =
    8\tau \Vert F - F_0 \Vert_{P_0}^2.
\end{align*}
Thus when \( v(F) = 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]} \) we have by
Lemma~\ref{lemma:norm}
\begin{equation*}
  \frac{v(F)}{R_0(F)}
  = 1.5 \frac{P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}}{P_0{[\bar{B}_{\tau,0}(F, \blank)]}}
  \leq 12 \tau,
\end{equation*}
and so using the Bernstein pairs \( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) we have
\begin{equation*}
  \sup_{F \in \Theta}
  \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
      \frac{1}{\delta} + 1
    \right)
  \right\}
  \leq \tau
  \left(
    13 + \frac{12}{\delta}
  \right),
\end{equation*}
For all $\delta>0$ we thus have
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad
    + (1+\delta)\log(1 + |\mathcal{F}_n|) \tau \frac{16 K}{n}
    \left(
    13 + \frac{12}{\delta}
    \right),
\end{align*}
and then the final result follows from Lemma~\ref{lemma:norm}.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:asymp-cons}]
  By definition of the oracle and Lemma~\ref{lemma:norm},
  \begin{equation*}
    \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]} \leq \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]}  
  \end{equation*}
  for all \( n \in \N \). The results then follows from
  Corollary~\ref{cor:oracle-prop}.
\end{proof}

>>>>

@@

\begin{corollary}
  \label{cor:oracle-prop}
  For all \(P_0\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    % \\
    % & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

@@

An example of a targeted estimator is the one-step estimator, defined as
\begin{equation}
    \label{eq:one-step-def}
    \hat{\Psi}_{\text{OS}}(\data_n)
    =
    \tilde{\Psi}^0(\hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
    H(\data_n))
    + \empmeas{[\psi(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
      \hat{b}_n(\data_n) )]},
\end{equation}
where \( \empmeas \) is the empirical measure of a data set
\(\{O_i\}_{i=1}^n\).
Under suitable regularity conditions
we have the
following asymptotic expansion of the one-step estimator
\citep{pfanzagl1985contributions,van2003unified,fisher2021visually,kennedy2022semiparametric},
\begin{equation*}
  \hat{\Psi}_{\text{OS}}(\data_n)- \Psi(P)
  =  \empmeas{[\psi(\blank ; \Lambda_1, \Lambda_2, \Gamma)]}
  +\mathrm{Rem}{(\hat{\Lambda}_{1,n},\hat{\Lambda}_{2,n},  \hat{\Gamma}_n, P)} + \smallO_{P}(n^{-1/2}),
\end{equation*}
where the remainder term has the form
\begin{equation}
  \label{eq:4}
  \mathrm{Rem}{(\hat{\Lambda}_{1,n},\hat{\Lambda}_{2,n},  \hat{\Gamma}_n, P)}
  = \mathcal{O}_P{
    \left\{
      \|\Lambda_1-\hat{\Lambda}_{1,n}\|^2
      +
      \|\Lambda_2-\hat{\Lambda}_{2,n}\|^2
      +
      \|\Gamma-\hat{\Gamma}_{n}\|^2
    \right\}
  },
\end{equation}
for some suitable norm \( \|\blank \| \), for instance the
\( \mathcal{L}_{P}^2 \)-norm. When equation~(\ref{eq:4}) holds and the
nuisance parameters $\Lambda_1$, $\Lambda_2$, and $\Gamma$ are
consistently estimated at rate \( \smallO_P{(n^{-1/4})} \), then
\begin{equation}
  \label{eq:3}
  \sqrt{n}(\hat{\Psi}_{\text{OS}}(\data_n)- \Psi(P)) \weakly \mathcal{N}(0,
  P{[\psi(\blank; \Lambda_1, \Lambda_2, \Gamma)^2]}),
\end{equation}
where we use \( \weakly \) to denote weak convergence \citep{van2000asymptotic}.
In particular, equation~(\ref{eq:3}) and Slutsky's lemma imply that we can
obtain asymptotically valid \((1-\alpha)\cdot100\%\) confidence intervals by
calculating
\begin{equation*}
  \left[
    \hat{\Psi}_{\text{OS}}(\data_n) - q_{\alpha/2} \hat{\sigma}(\data_n) ,
    \;
    \hat{\Psi}_{\text{OS}}(\data_n) + q_{\alpha/2} \hat{\sigma}(\data_n)
  \right],
\end{equation*}
where \( q_{\alpha} \) is the \( (1-\alpha) \)-quantile of the standard normal
distribution and
\( \hat{\sigma}(\data_n)^2 = \empmeas{ \left[ \psi(\blank;
    \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n), \hat{b}_n(\data_n))^2
  \right]} \).

@@

We compare the state learner to two IPCW based super learners: The
first super learner, called IPCW(Cox), uses a Cox model with main
effects to estimate the censoring probabilities, while the second
super learner, called IPCW(KM), uses the Kaplan-Meier estimator to
estimate the censoring probabilities. The Cox model for the censoring
distribution is correctly specified in both simulation settings while
the Kaplan Meier estimator only estimates the censoring model
correctly in the simulation setting where censoring is
independent. Both IPCW super learners are fitted using the
\texttt{R}-package \texttt{riskRegression}
\citep{Gerds_Ohlendorff_Ozenne_2023}.
%
% | time | sim_setting | true_events | true_cens | at_risk |
% |------+-------------+-------------+-----------+---------|
% |   36 | original    |      24.619 |    61.853 |  25.774 |
% |   36 | indep_cens  |      24.674 |    38.740 |  46.141 |
%
The IPCW super learners use the integrated Brier score up to a fixed time
horizon (36 months). The marginal risk of the event before this time horizon is
\(\approx 24.6\)\%. Under the ``dependent censoring'' setting the marginal
censoring probability before the time horizon is \(\approx 61.9\)\%. Under the
``independent censoring'' setting the marginal censoring probability before this
time horizon is \( \approx 38.7 \)\%.

@@

In this section we use the prostate cancer data of
\cite{kattan2000pretreatment} to illustrate the use of the state
learner in the presence of competing risks. We have introduced the
data in Section~\ref{sec:numer-exper}. The data consists of 1,042
patients who are followed from start of followup until tumor
recurrence, death without tumor recurrence or end of followup
(censored) whatever came first. For the sole purpose of illustration,
we estimate the average treatment effect of hormone therapy on death
and tumor recurrence. To do this we adapt the estimation strategy of
Section~\ref{sec:targeted-learning} as follows.  We use the state
learner to rank libraries of learners for the cause-specific
cumulative hazard functions of tumor recurrence, death without tumor
recurrence, and censoring.  The libraries of learners include the five
learners described in Table~\ref{tab:zel-library}.  We use the same
set of learners to learn the cumulative hazard function of tumor
recurrence \( \Lambda_1 \), the cumulative hazard function of death
without tumor recurrence \( \Lambda_2 \), and the cumulative hazard
function of the conditional censoring distribution $\Gamma$. 
We then use the highest ranked combination
of learners and apply formula~\eqref{eq:one-step-comp-ate}.

@@

\begin{figure}
  \centering %
  \includeFigCond[width=.9\linewidth]{real-data-state-learner.pdf}
  \caption[]{The results of applying the 125 combinations of learners to the
    prostate cancer data set. The learners are \texttt{KM} (KM), \texttt{Cox
      strata CS} (strata), \texttt{Lasso} (lasso), \texttt{Elastic} (elastic),
    and \texttt{RF} (RF) as described Table~\ref{tab:zel-library}. The error
    bars are based on five repetitions using different splits. We refer to
    learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as `Tumor
    learner', `Death learner', and `Censoring learner', respectively.}
  \label{fig:zelefski-real}
\end{figure}

\begin{table}
  \caption{\label{tab:zelefski-real}The 10 best performing models in terms of integrated Brier score. The
    reported standard errors are based on five repetitions using different
    splits. The models are described in Table~\ref{tab:zel-library}. We refer to
    learners of \( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as `Tumor
    learner', `Death learner', and `Censoring learner', respectively.}
  \centering
  \input{./real-data-tab.tex}
\end{table}

@@

\begin{figure}
  \centering%
  \includeFigCond[width=.8\linewidth]{real-data-target.pdf}
  \caption[]{Estimates of the average treatment effect of hormone therapy on the
    risk of tumor recurrence and death obtained from the prostate cancer study
    analyzed by \cite{kattan2000pretreatment}. The estimates are based on the
    estimator defined in equation~(\ref{eq:one-step-comp-ate}) with point wise
    95\% confidence intervals calculated as described in
    Section~\ref{sec:targeted-learning}. The estimates of the nuisance
    parameters are provided by the state learner.}
  \label{fig:zelefski-real-target}
\end{figure}

@@

\begin{proof}[Proof of Proposition~\ref{prop:stric-prop}.]
  The result follows from Lemma~\ref{lemma:norm}.
\end{proof}

Recall that $\mathcal{H}_{\mathcal{P}}$ denote the function space consisting of
all conditional state occupation probability functions for some measure
\( P \in \mathcal{P} \).

\begin{proof}[Proof of Corollary~\ref{cor:oracle-prop}.]
  First note that minimising the loss \( \bar{B}_{\tau} \) is equivalent to
  minimising the loss \( \bar{B}_{\tau,0} \), so the discrete super learner and
  oracle according to \( \bar{B}_{\tau} \) and \( \bar{B}_{\tau,0} \) are
  identical. By Lemma~\ref{lemma:norm}, \( R_0(F) \geq 0 \) for any
  \( F \in \mathcal{H}_{\mathcal{P}} \), and so using Theorem 2.3 from \citep{vaart2006oracle} with
  \( p=1 \), we have that for all \( \delta >0 \),
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad + (1+\delta) \frac{16 K}{n}
    \log(1 + |\mathcal{F}_n|)\sup_{F \in \mathcal{H}_{\mathcal{P}}}
    \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
    \frac{1}{\delta} + 1
    \right)
    \right\}
\end{align*}
where for each \( F \in \mathcal{H}_{\mathcal{P}} \), \( (M(F), v(F)) \) is some Bernstein pair for
the function \(o \mapsto \bar{B}_{\tau,0}(F, o) \). As
\( \bar{B}_{\tau,0}(F, \blank) \) is uniformly bounded by \( \tau \) for any
\( F \in \mathcal{H}_{\mathcal{P}} \), it follows from section 8.1 in \citep{vaart2006oracle} that
\( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) is a Bernstein pair for
\( \bar{B}_{\tau,0}(F, \blank) \). Now, for any \( a,b,c \in \R \) we have
\begin{align*}
  (a-c)^2 - (b-c)^2
  & = (a-b+b-c)^2 - (b-c)^2
  \\
  & = (a-b)^2 + (b-c)^2 +2(b-c)(a-b) - (b-c)^2
  \\
  & = (a-b)
    \left\{
    (a-b) +  2(b-c)
    \right\}
  \\
  & = (a-b)
    \left\{
     a + b -2c
    \right\},
\end{align*}
so using this with \( a=F(t, j, x) \), \( b=F_0(t, j, x) \), and
\( c = \1{\{\eta(t) = j\}} \), we have by Jensen's inequality
\begin{align*}
  & P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}
  \\
  & \leq
    2\tau\E_{P_0}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left\{
    \left(
    F(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    -
    \left(
    F_0(t, j, X) - \1{\{\eta(t) = j\}}
    \right)^2
    \right\}^2
    \diff t 
    \right]}
  \\
  & =2\tau
    \E_{P_0}\Bigg[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_0(t, j, X)
    \right)^2
  \\
  & \quad \quad \quad\quad \quad \quad \times
    \left\{
    F(t, j, X) +  F_0(t, j, X)-2 \1{\{\eta(t) = j\}}
    \right\}^2
    \diff t 
    \Bigg]
  \\
  & \leq
    8\tau \E_{P_0}{\left[
    \sum_{j=-1}^{2} \int_0^{\tau}
    \left(
    F(t, j, X) - F_0(t, j, X)
    \right)^2
    \diff t 
    \right]}.
  \\
  & =
    8\tau \Vert F - F_0 \Vert_{P_0}^2.
\end{align*}
Thus when \( v(F) = 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]} \) we have by
Lemma~\ref{lemma:norm}
\begin{equation*}
  \frac{v(F)}{R_0(F)}
  = 1.5 \frac{P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}}{P_0{[\bar{B}_{\tau,0}(F, \blank)]}}
  \leq 12 \tau,
\end{equation*}
and so using the Bernstein pairs \( (\tau, 1.5 P_0{[\bar{B}_{\tau,0}(F, \blank)^2]}) \) we have
\begin{equation*}
  \sup_{F \in \mathcal{H}_{\mathcal{P}}}
  \left\{
    M(F) + \frac{v(F)}{R_0(F)}
    \left(
      \frac{1}{\delta} + 1
    \right)
  \right\}
  \leq \tau
  \left(
    13 + \frac{12}{\delta}
  \right),
\end{equation*}
For all $\delta>0$ we thus have
\begin{align*}
  \E_{P_0}{\left[ R_0(\hat{\phi}_n(\data_n^{-k})) \right]}
  \leq
  &(1+2\delta)\E_{P_0}{\left[ R_0(\tilde{\phi}_n(\data_n^{-k})) \right]}
  \\
  & \quad
    + (1+\delta)\log(1 + |\mathcal{F}_n|) \tau \frac{16 K}{n}
    \left(
    13 + \frac{12}{\delta}
    \right),
\end{align*}
and then the final result follows from Lemma~\ref{lemma:norm}.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:asymp-cons}.]
  By definition of the oracle and Lemma~\ref{lemma:norm},
  \begin{equation*}
    \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]} \leq \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]}  
  \end{equation*}
  for all \( n \in \N \). The results then follows from
  Corollary~\ref{cor:oracle-prop}.
\end{proof}
